---
title: "Mini Literate Analysis"
author: "Jacob Schmidt"
date: 10/02/2023
format: html
---

## Introduction

### About this data
The dataset used for this work includes chemical and physical parameters from sea ice platforms in the Arctic in order to provide context for how high latitude oceans regulate global climate. Data was collected in April or May of 2000-2014 and has been deposited in the [Arctic Data Center](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X)

### Setup

```{r}
#| warning: FALSE
library(readr)
library(here)
```

### Read data

```{r}
#| warning: FALSE
bg_chem <- read_csv(here("data/BGchem2008data.csv"))
```
Exploring our data: 
```{r}
colnames(bg_chem)
str(bg_chem)
summary(bg_chem)
unique(bg_chem$Date)
```


## Analysis

### Calculate summary statistics

```{r}
nitrate <- mean(bg_chem$NO3)
nitrite <- mean(bg_chem$NO2)
ammonium <- mean(bg_chem$NH4)
phosphorous <- mean(bg_chem$P)
```

### Calculate mean Redfield Ratio
```{r}
RR <- (nitrate + nitrite + ammonium) / phosphorous
RR
```

### Plot Redfield Ratio
```{r}
plot(bg_chem$P, bg_chem$NO2 + bg_chem$NO3 + bg_chem$NH4)
```


## Conclusion

The Redfield Ratio of this data is `r round(RR, digits = 3)`