---
title: "Cleaning and Wrangling Data"
format: html
---

## Set up

```{r}
#| warning: FALSE
#| message: FALSE

library(dplyr)
library(tidyr)
library(readr)
```

## Read data

```{r}
#| warning: FALSE
#| message: FALSE

catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
```

```{r}
#| echo: FALSE
#| eval: FALSE

colnames(catch_original)
summary(catch_original)
head(catch_original)
unique(catch_original$Region)
```

## Use `select()` to remove or select columns.

The following code chunks function the same. The (-) removes the subsequent columns.

```{r}
catch_data <- catch_original %>%
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)

catch_data <- catch_original %>%
  select(-All, -notesRegCode)
```

Quality check:

```{r}
glimpse(catch_data)
```

## Overwrite Chinook column using `mutate()`.

We want the Chinook values to be <dbl>, not <chr>.

```{r}
catch_clean <- catch_data %>%
  mutate(Chinook = as.numeric(Chinook))
```

```{r}

# Base R was used to identify which column had the problematic character value of the Chinook column.
i <- which(is.na(catch_clean$Chinook))
i
catch_data[i,]
```

```{r}
catch_clean <- catch_data %>%
  mutate(Chinook = if_else(condition = Chinook == "I",
                           true = "1",
                           false = Chinook),
         Chinook = as.numeric(Chinook))

glimpse(catch_clean)
```

## Reshaping DF with `pivot_longer()` and `pivot_wider()`.

```{r}
catch_long <- catch_clean %>%
  pivot_longer(cols = -c(Region, Year), 
               names_to = "Species", 
               values_to = "Catch")
head(catch_long)
```

## The `rename()` function.
```{r}
# catch_long <- catch_long %>%
#   rename(Catch_Thousands = Catch_Thousand)
# head(catch_long)
```

## Adding a column using `mutate()`.
```{r}
catch_long <- catch_long %>%
  mutate(Catch = Catch_Thousands * 1000) %>%
  select(-Catch_Thousands)

head(catch_long)
```

## Split, Apply, Combine
```{r}
mean_region <- catch_long %>%
  group_by(Region) %>%
  summarise(mean_catch = mean(Catch))

head(mean_region)

n_region <- catch_long %>%
  group_by(Region) %>%
  summarize(n = n())

head(n_region)
```

## Filtering using `filter()`.
```{r}
sse_catch <- catch_long %>%
  filter(Region == "SSE")

sse_catch_2 <- catch_long %>%
  filter(Region %in% c("SSE", "NSE"))
```

